### 2017-10-12 manuscript.Rmd
Initial commit of local files

# Question based metrics

## Introduction

Scholarly research has been said to face a reproducibility crisis, but the underlying cause of this is the incentive system.

### Current metrics

1. backward looking
2. invalid
    * journal impact factor
    * citations
    * h-index
3. not question based
    * use bad proxy because nothing else is available ("we have to do something")

## Granular communication

### Network structure

d

## Renewing metrics

## Use cases

### Funders

### Individual researchers

### Evaluation committees

### Universities

## Discussion